\section{Conclusion}
\label{sec:conclusion}
% xmlmate is a good defect search tool with extensions presented here
\xmlmate is a good search-based tool for finding defects that induce crashes in the tested program. It has
become especially versatile after being enhanced with the extensions proposed in this document.
% list extensions: support for binaries, pluggable fitfuncs, additional formats, schema violations
The part of the genetic algorithm responsible for fitness score representation has profited from adding
support for both pluggable and vulnerability-specific fitness functions. Through this new interface \xmlmate
has gained the ability to produce fitness scores for arbitrary binaries as it is no longer limited to
the \java Virtual Machine, on which it is running itself. By leveraging various file format specifications and
providing converters for those formats, it has now become possible to generate inputs compatible with a wide
variety of potential test subjects previously inaccessible. For purposes of security testing the main
functionality of the input generation engine has been extended to include intentional specification deviations.

% equipped with these extensions xmlmate yielded good results in identifying defects
% its efficiency and effectiveness are justified by it finding known vulns a lot of times over
Equipped with these additions, \xmlmate yielded excellent results in an experiment designed to show its
effectiveness and efficiency as it was able to repeatedly identify known vulnerabilities in a ubiquitously
used library, while maintaining a zero false positives rate by design.
% its practical use is strengthened by it finding unknown bugs even when not specifically targeting
Its practicability is further emphasized by the fact that it has found defects previously unknown in programs
which were not even specifically targeted in the experiments. 
% quality of results (fitness scores and crashes) depends on quality of test drivers and schema and converters
It must be also noted, however, that the results are highly dependent on the quality of the user provided
schema and test harness.

% overall xmlmate shows potential
\begin{mdframed}
Overall, \xmlmate demonstrated great value as a format-aware search-based defect detection tool which
is effective, efficient, and practical.
\end{mdframed}